\bf\mftitle

### 2011-01-10 doc/jInfer/full/architecture/architecture.tex
3154619 Create documentation for jInfer architecture - little folder #2

For general information on regular expressions, please refer to \cite{wikiregexp}, \cite{Hopcroft:2006:IAT:1177300}.

All classes pertaining to regular expressions can be found in the package \code{cz.cuni.mff.ksi.jinfer.base.regexp}.

In theory, regular expression is defined recursively:
\begin{itemize}
	\item Every symbol $x$ from alphabet is regular expression representing language ${x}$.
	\item If $R_1, R_2, \dots, R_n$ are regular expressions representing languages $L_1, L_2, \dots, L_n$, then $R_1 | R_2 | \dots | R_n$
	is regular expression representing language $\bigcup_{i = 1}^{n}{L_i}$. Operation $|$ is called alternation.
	\item If $R_1, R_2, \dots, R_n$ are regular expressions representing languages $L_1, L_2, \dots, L_n$, then $R_1, R_2 , \dots , R_n$
	is regular expression representing language $L_1 \cdot L_2 \cdot \ldots \cdot L_n$. Where $\cdot$ is languague concatenation,
	and operation $,$ is called concatenation as well.
	\item If $R$ is regular expression representing language $L$, then $R*$ 
	is regular expression representing language $L*$. In regular expressions, operator $*$ is called Kleene star.
\end{itemize}

In common practice, extended regular expressions are commonly used. Although they are not extending set of
languages representable by regular expressions, they give us nicer syntax. Extended regular expression adds syntax:
\begin{itemize}
	\item $R{m,n}$, where $m,n \in N$ with meaning of $R$ at least $m$-times, at most $n$-times,
	\item $R{m,}$ with meaning at least $m$-times,
	\item $R?$ with meaning $R{0,1}$,
	\item $R+$ with meaning $R{1,}$.
\end{itemize}

By this syntax, Kleene star is simply interval ${0,}$. This is just syntactic shortcut for long expresssions like $R, R, R, R?, R?$, which is just $R{3,5}$.

In jInfer, we implement regular expresssions as class \code{Regexp} with supplementing classes \code{RegexpInterval} and \code{RegexpType}.
Each \code{Regexp} instance has one of the \code{RegexpType} type:
\begin{itemize}
	\item Token - a letter of the language.
	\item Concatenation - one or more regular expression in an ordered sequence. Eg. $(a, b, c, d)$.
	\item Alternation - a choice between one or more regular expressions. Eg. $(a | b | c | d)$.
	\item Permutation - shortcut for all possible permutations of regular expressions (to work with \code{xs:all} in future).
	Our syntax to write down permutation is $(a  b c  d)$.
\end{itemize}

Each \code{Regexp} instance has interval associated with it (as a member). Class \code{RegexpInterval} represents all kinds of intervals
${m,n}; {m,}$.

>>>>

For general information on regular expressions, please refer to \cite{wikiregexp}, \cite{automatatheory}.
All classes pertaining to regular expressions can be found in the package \code{cz.cuni.mff.ksi.jinfer.base.regexp}.
In jInfer, we use extended regular expressions as they give us nicer syntax (and easier programming).

Regular expression is implemented as class \code{Regexp} with supplementing classes \code{RegexpInterval} and \code{RegexpType}.
Each \code{Regexp} instance has one of the enum \code{RegexpType} type:
\begin{itemize}
	\item Lambda - empty string (also called $\epsilon$ in literature),
	\item Token - a letter of the alphabet,
	\item Concatenation - one or more regular expression in an ordered sequence. Eg. $(a, b, c, d)$,
	\item Alternation - a choice between one or more regular expressions. Eg. $(a | b | c | d)$,
	\item Permutation - shortcut for all possible permutations of regular expressions. Our syntax to write down permutation is $(a\& b\& c\& d)$.
\end{itemize}

Each \code{Regexp} instance has one instance of \code{RegexpInterval} as member.
Class \code{RegexpInterval} represents POSIX-like intervals for expression:
\begin{itemize}
	\item $a\{m,n\}$ means $a$ at least $m$-times, at most $n$-times,
	\item $a\{m,\}$ means at least $m$-times (unbounded).
\end{itemize}

Class \code{Regexp} can represent regular expression over any alphabet.
This is done by using java generics, \code{Regexp} is implemented as \code{Regexp<T>}.
Regular expression is in fact $n$-ary tree, for example expression $(a, b, (c | d))$ can be viewed as in figure \ref{reg:tree}.
\begin{figure}
\caption{Example tree for regular expression $(a, b, (c | d))$} \label{reg:tree}
\includegraphics[scale=1]{reg_tree}
\end{figure}
